---
title: "Bayesian Meta Analyses"
author: "Candice"
date: "2024-06-17"
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Separate Bayesian Meta Analysese for Each Outcome

This file constructs the Bayesian Meta Analysis for each of the five outcomes separately, namely: infection, symptomatic infection, severe infection, hospitalisation and death. We have conducted prior predictive checks to assess the fit of the chosen prior distributions to the observed data. There are three choices: non-informative priors, weakly informative priors with half-Cauchy, weakly informative priors with half-normal.


```{r, warning = FALSE, message = FALSE}
# Library relevant packages
library(brms)
library(metafor) 
library(rstan) 
library(kableExtra) 
library(ggplot2) 
library(bayesplot)
library(gridExtra) 
library(ggplotify) 
library(ggridges) 
library(tidybayes) 
library(dplyr) 
library(glue) 
library(stringr) 
library(forcats)
library(posterior)

color_scheme_set("red")

# Load data for each of the five outcomes
# with the first row containing variable names 
infec.data <- read.csv("infection_data.csv", header = TRUE)
sym.infec.data <- read.csv("symptomatic_infection_data.csv", header = TRUE)
sev.infec.data <- read.csv("severe_infection_data.csv", header = TRUE)
hos.data <- read.csv("hospitalisation_data.csv", header = TRUE)
death.data <- read.csv("death_data.csv", header = TRUE)

# Load helper functions 
# Supplementary file with R functions
source("supplementary code functions.R")
```


```{r, warning = FALSE, message = FALSE}
## Helper function to prepare data for Bayesian meta-analysis
# Calculate log relative risks, relative risks, variances, and SEs
prepare_data <- function(data){
  es <- escalc(measure = "RR", 
               ai = n1, n1i = N1, 
               ci = n2, n2i = N2, 
               slab = author_id, data = data)
  es$RR <- exp(es$yi)
  es$sei <- sqrt(es$vi)
  
  # Remove the third and fourth columns corresponding to intervention indication variables
  es <- es[, -c(3, 4)]
  
  return(es)
}

# Prepare data for each outcome
infec.es <- prepare_data(infec.data)
sym.infec.es <- prepare_data(sym.infec.data)
sev.infec.es <- prepare_data(sev.infec.data)
hos.es <- prepare_data(hos.data)
death.es <- prepare_data(death.data)
```


## Prior Predictive Checks

```{r, warning = FALSE, message = FALSE}
## Define a function to run Bayesian meta-analysis with prior predictive check
run_prior_predictive_checks <- function(data, model_name, priors, seed = 18006513){
  # Bayesian meta-analysis model using brms package
  model <- brm(
    data = data,
    family = "gaussian",
    yi | se(sei) ~ 1 + (1 | author_id),
    prior = priors,
    iter = 2000, warmup = 1000, cores = 4, chains = 4,
    control = list(adapt_delta = 0.99),
    # Perform prior predictive checks
    sample_prior = "only", 
    seed = seed
  )
  
  # Prior predictive check plot
  plot <- pp_check(model, ndraws = 100)
  
  # Store model and plot in global environment
  assign(paste0(model_name, ".prior"), model, envir = .GlobalEnv)
  assign(paste0(model_name, ".priorpc.plot"), plot, envir = .GlobalEnv)
}


## Define priors for different models
# Model 1: Non-informative Priors
priors_non_informative <- c(prior(normal(0, 10000), class = Intercept), 
                            prior(uniform(0, 10000), class = sd))

# Model 2: Weakly Informative Priors with Half-Cauchy
priors_half_cauchy <- c(prior(normal(0, 1), class = Intercept), 
                        prior(cauchy(0, 0.5), class = sd))

# Model 3: Weakly Informative Priors with Half-Normal
priors_half_normal <- c(prior(normal(0, 1), class = Intercept), 
                        prior(normal(0, 0.5), class = sd, lb = 0))

```


```{r, warning = FALSE, message = FALSE, fig.cap = "Model Prior Predictive Checks (Infection)"}
## Run prior predictive checks for each outcome and prior model

# Infection
run_prior_predictive_checks(infec.es, "infec1", priors_non_informative)
run_prior_predictive_checks(infec.es, "infec2", priors_half_cauchy)
run_prior_predictive_checks(infec.es, "infec3", priors_half_normal)

# Plot results for infection models
infec1.priorpc.plot
infec2.priorpc.plot
infec3.priorpc.plot
```


```{r, warning = FALSE, message = FALSE, fig.cap = "Model Prior Predictive Checks (Symptomatic Infection)"}
# Symptomatic Infection
run_prior_predictive_checks(sym.infec.es, "sym1", priors_non_informative)
run_prior_predictive_checks(sym.infec.es, "sym2", priors_half_cauchy)
run_prior_predictive_checks(sym.infec.es, "sym3", priors_half_normal)

# Plot results for symptomatic infection models
sym1.priorpc.plot
sym2.priorpc.plot
sym3.priorpc.plot
```


```{r, warning = FALSE, message = FALSE, fig.cap = "Model Prior Predictive Checks (Severe Infection)"}
# Severe Infection
run_prior_predictive_checks(sev.infec.es, "sev1", priors_non_informative)
run_prior_predictive_checks(sev.infec.es, "sev2", priors_half_cauchy)
run_prior_predictive_checks(sev.infec.es, "sev3", priors_half_normal)

# Plot results for severe infection models
sev1.priorpc.plot
sev2.priorpc.plot
sev3.priorpc.plot
```


```{r, warning = FALSE, message = FALSE, fig.cap = "Model Prior Predictive Checks (Hospitalisation)"}
# Hospitalisation
run_prior_predictive_checks(hos.es, "hos1", priors_non_informative)
run_prior_predictive_checks(hos.es, "hos2", priors_half_cauchy)
run_prior_predictive_checks(hos.es, "hos3", priors_half_normal)

# Plot results for hospitalisation models
hos1.priorpc.plot
hos2.priorpc.plot
hos3.priorpc.plot
```


```{r, warning = FALSE, message = FALSE, fig.cap = "Model Prior Predictive Checks (Death)"}
# Death
run_prior_predictive_checks(death.es, "death1", priors_non_informative)
run_prior_predictive_checks(death.es, "death2", priors_half_cauchy)
run_prior_predictive_checks(death.es, "death3", priors_half_normal)

# Plot results for death models
death1.priorpc.plot
death2.priorpc.plot
death3.priorpc.plot
```



## Posterior Predictive Checks

```{r, warning = FALSE, message = FALSE}
# Set seed for reproducibility
set.seed(18006513)

## Define function to prepare trace plots and calculate R-hats statistics
plot_and_rhat <- function(model, params){
    # Prepare posterior array
    posterior_array <- as.array(model)
    dimnames(posterior_array)$variable <- c("Intercept", "Tau", "sigma", "Intercept2",
                                            params, "lprior", "lp__")
    
    # Trace plots
    trace_plot <- mcmc_trace(posterior_array, pars = c("Intercept", "Tau", params))
    print(trace_plot)
    
    # Calculate R-hats
    print(rhat(model))
}


## Define function for posterior predictive checks
posterior_predictive_check <- function(model, ndraws) {
    pp_check_plot <- pp_check(model, ndraws = ndraws)
    print(pp_check_plot)
}
```

### Infection

```{r, warning = FALSE, message = FALSE}
### Bayesian model estimation for each outcome ###

## Infection
for (i in 1:3) {
    # Define priors based on model type
    priors <- switch(i,
                     priors_non_informative,
                     priors_half_cauchy,
                     priors_half_normal)
    
    # Run Bayesian model
    infec_mod <- brm(data = infec.es,
                     family = "gaussian",
                     yi | se(sei) ~ 1 + (1 | author_id),
                     prior = priors,
                     iter = 2000, warmup = 1000, cores = 4, chains = 4,
                     control = list(adapt_delta = 0.99),
                     seed = 18006513)
    
    # Print model summary
    print(summary(infec_mod))
    
    # Plot trace and calculate R-hats
    plot_and_rhat(infec_mod, paste("RE:", 1:15))
    
    # Perform posterior predictive check
    posterior_predictive_check(infec_mod, ndraws = 200)
}
```

### Symptomatic Infection

```{r, warning = FALSE, message = FALSE}
## Symptomatic Infection
for (i in 1:3) {
    priors <- switch(i,
                     priors_non_informative,
                     priors_half_cauchy,
                     priors_half_normal)
    
    # Run Bayesian model
    sym_infec_mod <- brm(data = sym.infec.es,
                         family = "gaussian",
                         yi | se(sei) ~ 1 + (1 | author_id),
                         prior = priors,
                         iter = 2000, warmup = 1000, cores = 4, chains = 4,
                         control = list(adapt_delta = 0.99),
                         seed = 18006513)
    
    # Print model summary
    print(summary(sym_infec_mod))
    
    # Plot trace and calculate R-hats
    plot_and_rhat(sym_infec_mod, paste("RE:", 1:5))
    
    # Perform posterior predictive check
    posterior_predictive_check(sym_infec_mod, ndraws = 200)
}
```

### Severe Infection

```{r, warning = FALSE, message = FALSE}
## Severe Infection
for (i in 1:3) {
    priors <- switch(i,
                     priors_non_informative,
                     priors_half_cauchy,
                     priors_half_normal)
    
    # Run Bayesian model
    sev_infec_mod <- brm(data = sev.infec.es,
                         family = "gaussian",
                         yi | se(sei) ~ 1 + (1 | author_id),
                         prior = priors,
                         iter = 2000, warmup = 1000, cores = 4, chains = 4,
                         control = list(adapt_delta = 0.99),
                         seed = 18006513)
    
    # Print model summary
    print(summary(sev_infec_mod))
    
    # Plot trace and calculate R-hats
    plot_and_rhat(sev_infec_mod, paste("RE:", 1:6))
    
    # Perform posterior predictive check
    posterior_predictive_check(sev_infec_mod, ndraws = 200)
}
```

### Hospitalisation

```{r, warning = FALSE, message = FALSE}
## Hospitalisation
for (i in 1:3) {
    priors <- switch(i,
                     priors_non_informative,
                     priors_half_cauchy,
                     priors_half_normal)
    
    # Run Bayesian model
    hos_mod <- brm(data = hos.es,
                   family = "gaussian",
                   yi | se(sei) ~ 1 + (1 | author_id),
                   prior = priors,
                   iter = 2000, warmup = 1000, cores = 4, chains = 4,
                   control = list(adapt_delta = 0.99),
                   seed = 18006513)
    
    # Print model summary
    print(summary(hos_mod))
    
    # Plot trace and calculate R-hats
    plot_and_rhat(hos_mod, paste("RE:", 1:4))
    
    # Perform posterior predictive check
    posterior_predictive_check(hos_mod, ndraws = 200)
}
```


### Death

```{r, warning = FALSE, message = FALSE}
## Death
for (i in 1:3) {
    priors <- switch(i,
                     priors_non_informative,
                     priors_half_cauchy,
                     priors_half_normal)
    
    # Run Bayesian model
    death_mod <- brm(data = death.es,
                     family = "gaussian",
                     yi | se(sei) ~ 1 + (1 | author_id),
                     prior = priors,
                     iter = 2000, warmup = 1000, cores = 4, chains = 4,
                     control = list(adapt_delta = 0.99),
                     seed = 18006513)
    
    # Print model summary
    print(summary(death_mod))
    
    # Plot trace and calculate R-hats
    plot_and_rhat(death_mod, paste("RE:", 1:3))
    
    # Perform posterior predictive check
    posterior_predictive_check(death_mod, ndraws = 200)
}
```

